---
title: "Untitled"
author: "Davide Risso"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE)
```


# Workflow

1. filter high variance bw replicates of same condition
2. Correlation bw average profiles across conditions
3. Rank the proteins by correlation (possibly divided by variance)

```{r datain}
library(magrittr)
library(ggplot2)
library(DepLabData)
library(DepLab)
library(tidyverse)
library(matrixStats)

data("WT_trial1")
data("WT_trial2")
data("DN_trial1")
data("DN_trial2")
data("EV_trial1")
data("EV_trial2")

WT_trial1 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> wild_type_1
WT_trial2 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> wild_type_2

DN_trial1 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> dn_1
DN_trial2 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> dn_2

EV_trial1 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> ev_1
EV_trial2 %>%
  MQ_to_longFormat(., y = "raw.intensity", return.dt = TRUE,
                   extract_proteinID(.$Protein.IDs, routine = "human")) -> ev_2

## combine samples
wild_type_1 %>%
  mutate(expt_id = "wt1", exp_cond = "wt") -> wt1
wild_type_2 %>%
  mutate(expt_id = "wt2", exp_cond = "wt") -> wt2
dn_1 %>%
  mutate(expt_id = "dn1", exp_cond = "dn") -> dn1
dn_2 %>%
  mutate(expt_id = "dn2", exp_cond = "dn") -> dn2
ev_1 %>%
  mutate(expt_id = "ev1", exp_cond = "ev") -> ev1
ev_2 %>%
  mutate(expt_id = "ev2", exp_cond = "ev") -> ev2

long_df <- rbind(wt1, wt2, dn1, dn2, ev1, ev2)

## smoothing
smu <- superSmooth_values(long.df = long_df, prot.identifier = "id")

## smoothing will create negative values --> putting them to 0
smu %>%
  mutate(value = pmax(0, value)) -> smu

## wide data
wide_smu <- na.omit(dcast(smu, "id + fraction ~ expt_id"))
```

# Filtering

For now we start from all the data, but in the app the starting point will be already filtered for protein quality. We start from smoothed data.

Nonetheless, we need some more sophisticated filtering here, i.e., filtering out proteins that have high variance between replicates of the same condition. In case of more than one condition, we want to be strict and remove a protein if it has high variance in any condition.

We start by looking at the correlation between replicates of the EV condition.

```{r filtering}
## first, filter out those proteins that are not detected in all replicates
protein_list <- unique(as.character(wide_smu$id))
length(protein_list)

smu %>%
  filter(id %in% protein_list) -> smu_filtered

## compute correlation between replicates for each condition
smu_filtered %>%
  filter(exp_cond == "ev") %>%
  group_by(id) %>%
  do(data.frame(cor = cor(log1p(.$value[.$expt_id == "ev1"]),
                       log1p(.$value[.$expt_id == "ev2"]))))

## example of good correlation
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A1L0T0") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of good correlation")

## example of bad correlation 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A1L390") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of bad correlation")

## example of so-so correlation 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A0A0B4J2F0") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of so-so correlation")
```

To generalize to more than two samples, we can use variance.

```{r variance}
## compute fraction by fraction variance
smu_filtered %>%
  filter(exp_cond == "ev") %>%
  dcast("id + fraction ~ expt_id") %>%
  mutate(cv = rowSds(cbind(log1p(ev1), log1p(ev2))) / 
               rowMeans(cbind(log1p(ev1), log1p(ev2)))) %>%
  group_by(id) %>%
  summarize(mean_cv = mean(cv, na.rm = TRUE))
  
## example of large variance 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A0JLT2") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of large variance")
  
## example of small variance 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A1L390") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of small variance")
```

So variance / coefficient of variation is a bad idea. To generalize the correlation, we can either do all pairwise correlations or correlations to the mean profile. And than get the minimum correlation (too strict?).

To test out these ideas, I will pretend that dn and wt are four replicates of the same condition.

To be more robust to really large values, we compute all the correlations on the log scale.

```{r four_replicates}
smu_filtered %>%
  mutate(value = log1p(value)) -> smu_log

smu_log %>%
  filter(exp_cond %in% c("dn", "wt")) %>%
  dcast("id + fraction ~ expt_id") %>%
  group_by(id) %>%
  do(data.frame(min(cor(.[,-(1:2)]))))

## example of good correlation
smu_filtered %>%
  filter(exp_cond %in% c("dn", "wt") & id == "A0FGR8") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of good correlation")

## example of bad correlation 
smu_filtered %>%
  filter(exp_cond%in% c("dn", "wt") & id == "A1L390") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of bad correlation")

## example of so-so correlation 
smu_filtered %>%
  filter(exp_cond %in% c("dn", "wt") & id == "A0A0B4J2F0") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of so-so correlation")
```

Final test: let's try the general method for two replicates.

```{r two_replicates}
smu_log %>%
  filter(exp_cond == "ev") %>%
  dcast("id + fraction ~ expt_id") %>%
  group_by(id) %>%
  do(data.frame(min(cor(.[,-(1:2)]))))

## example of good correlation
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A1L0T0") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of good correlation")

## example of bad correlation 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A1L390") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of bad correlation")

## example of so-so correlation 
smu_filtered %>%
  filter(exp_cond == "ev" & id == "A0FGR8") %>%
  ggplot(aes(x = fraction, y = value, color = expt_id)) + 
  geom_line() + ggtitle("example of so-so correlation")
```

Now we can look at the correlation distribution and see if there is a good threshold.

```{r cor_distro}
conditions <- c("ev", "wt", "dn")
correlations <- list()
correlations <- lapply(conditions, function(cond) {
  smu_log %>%
    filter(exp_cond == cond) %>%
    dcast("id + fraction ~ expt_id") %>%
    group_by(id) %>%
    do(data.frame(min_cor =  min(cor(.[,-(1:2)])))) %>%
    mutate(exp_cond = cond)
})

correlations <- do.call(rbind, correlations)

ggplot(correlations, aes(x = exp_cond, y = min_cor)) +
  geom_boxplot()

correlations %>%
  group_by(exp_cond) %>%
  summarize(q25 = quantile(min_cor, probs = .25, na.rm=TRUE))
```

Based on the overall distribution .5 seems a good threshold, but this is something that will likely change between datasets, so I would let the user decide. We filter out all the proteins that have a minimum correlation of less than .8 in at least one condition.

```{r cor_filter}
correlations %>%
  group_by(id) %>%
  summarize(cor = min(min_cor)) %>%
  filter(cor >= .5) %>%
  pull(id) %>% as.character -> good_proteins

smu_filtered %>%
  filter(id %in% good_proteins) -> smu_good
```

`smu_good` contains the proteins that we want to include in the analysis.

# Correlation

Now that we have a set of proteins that are consistently measured across replicates, we can look at correlation across average profiles.

We should compute all three pairwise comparisons. This might be done more efficiently, I guess, but for now just a for loop.

```{r all_conditions}
smu_good %>%
  group_by(id, exp_cond, fraction) %>%
  summarize(mean_value = mean(value)) -> mean_profiles

mean_profiles %>%
  mutate(mean_value = log1p(mean_value)) -> mean_profiles_log

comparisons <- list("EVvDN" = c("ev", "dn"),
                    "EVvWT" = c("ev", "wt"),
                    "WTvDN" = c("wt", "dn"))

bw_group_cors <- lapply(comparisons, function(comp) {
  mean_profiles_log %>%
    filter(exp_cond %in% comp) %>%
    dcast("id + fraction ~ exp_cond") %>%
    group_by(id) %>%
    set_colnames(c("id", "fraction", "cond1", "cond2")) %>%
    do(data.frame(cor = cor(.$cond1, .$cond2))) %>%
    arrange(cor)
})
bw_group_cors


mean_profiles %>%
  filter(id == "Q93063") %>%
  ggplot(aes(x = fraction, y = mean_value, color = exp_cond)) + 
  geom_line()

smu_good %>%
  filter(id == "Q93063") %>%
  ggplot(aes(x = fraction, y = value, color = exp_cond, 
             group=expt_id)) + 
  geom_line()

correlations %>% filter(id == "Q93063")

mean_profiles %>%
  filter(id == "P22234") %>%
  ggplot(aes(x = fraction, y = mean_value, color = exp_cond)) + 
  geom_line()

smu_good %>%
  filter(id == "P22234") %>%
  ggplot(aes(x = fraction, y = value, color = exp_cond, 
             group=expt_id)) + 
  geom_line()

correlations %>% filter(id == "P22234")

mean_profiles %>%
  filter(id == "Q8N257") %>%
  ggplot(aes(x = fraction, y = mean_value, color = exp_cond)) + 
  geom_line()

smu_good %>%
  filter(id == "Q8N257") %>%
  ggplot(aes(x = fraction, y = value, color = exp_cond, 
             group=expt_id)) + 
  geom_line()

correlations %>% filter(id == "Q8N257")
```

# Visualize results

```{r visualize}
bw_group_cors[[1]] %>%
  pull(id) %>% as.character %>% head(100) -> ids

mean_profiles %>%
  filter(id %in% ids & exp_cond == "dn") %>%
  dcast("id ~ fraction") -> dn_mat

mean_profiles %>%
  filter(id %in% ids & exp_cond == "wt") %>%
  dcast("id ~ fraction") -> wt_mat

mean_profiles %>%
  filter(id %in% ids & exp_cond == "ev") %>%
  dcast("id ~ fraction") -> ev_mat

stopifnot(all(ev_mat[,1] == wt_mat[,1]))
stopifnot(all(ev_mat[,1] == dn_mat[,1]))

mat <- as.matrix(cbind(ev_mat[,-1], wt_mat[,-1], dn_mat[,-1]))
colnames(mat) <- c(paste0("ev", 1:30), paste0("wt", 1:30),
                   paste0("dn", 1:30))
rownames(mat) <- ev_mat[,1]

NMF::aheatmap(log1p(mat), Colv = NA, distfun = "pearson",
         scale = "none")
```

# Session Info

```{r}
sessionInfo()
```
