---
title: "Running and testing the app"
author: "ABC @ WCM"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
---
<!--
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Testing correlation heatmaps}
  %\VignetteEncoding{UTF-8}
-->

```{r echo = FALSE}
library(knitr)
opts_chunk$set(cache = FALSE, echo = TRUE, message = FALSE, warning = FALSE, eval = FALSE)
library(ggplot2)
```

DepLab is now available as a package.
All the `library()` and `source()` calls have been disabled (except for the app's functions).

To **load the package**:

1. Make sure you have `devtools` installed.
2. Set the working directory to "DepLab".
3. Type `devtools::load_all()`.

To start the app, type `runPCP()`.

To check a specific function, you can now simply type `?MQ_to_longFormat()`.
This will show the documentation in the _Help_ panel of RStudio.

Once the package is loaded, you can access the example data within the `inst/` 
folder via `system.file(<FOLDER TO LOOK INTO>, <SUBFOLDER TO LOOK INTO>, <FILE NAME>, package = "DepLab")`.
This will return the full, absolute path to the file on your system.

### Testing yeast routine for generating long data.frame

```{r yeastRoutine}
devtools::load_all()

mq1 <- reading_MQ(system.file("extdata","test_data", "proteinGroups_100mM_new.txt", package = "DepLab"))
mqclean <- cleaning_MQ(mq1, remove.contaminants = TRUE, remove.decoys = TRUE, poi = "yeast")
ids.list <- extract_proteinID(mqclean$Protein.IDs, routine = "yeast")
mq.raw.intensity <- MQ_to_longFormat(mqclean, y = "raw.intensity", return.dt = TRUE, ids.list)
head(mq.raw.intensity)
```

### Reading in spike-ins

```{r spikeInRoutine}
std <- cleaning_MQ(mq.df = mq1, remove.contaminants = FALSE, remove.decoys = TRUE,
                   poi = NULL, spikeIn = c("P07477", "Q0140","YAL003W"))                                          
stdid <- extract_proteinID(std$Protein.IDs, regex = c(".*(P07477|Q0140|YAL003W).*", "\\1"), label = "id", routine = NULL)
std.raw.intensity <- MQ_to_longFormat(std, y = "raw.intensity", return.dt = TRUE, stdid)
```

### Testing human routine for generating long data.frame

```{r humanRoutine}
devtools::load_all()

mq.h <- reading_MQ(system.file("extdata","test_data", "proteinGroups_human.txt", package = "DepLab"))
mq.h.clean <- cleaning_MQ(mq.h, remove.contaminants = TRUE, remove.decoys = TRUE, poi = "human")
ids.list <- extract_proteinID(mq.h.clean$Protein.IDs, routine = "human")
mq.h.raw.intensity <- MQ_to_longFormat(mq.h.clean, y = "raw.intensity", return.dt = FALSE,ids.list)
head(mq.h.raw.intensity)
```

### Testing human routine for newer version of MaxQuant output

```{r newMQformat}
mq.h.new <- reading_MQ("~/Documents/Projects/2015-11_NoahDephoure_Proteomics/fromNoah_2016-03/proteinGroups_MCF10A_dounce_TSK_new.txt")
mq.h.new.clean <- cleaning_MQ(mq.h.new, remove.contaminants = TRUE, remove.decoys = TRUE, poi = "human")
ids.h.new <- extract_proteinID(mq.h.new.clean$Protein.IDs, routine = "human")
mq.h.raw.intensity <- MQ_to_longFormat(mq.h.new.clean, y= "raw.intensity", return.dt = TRUE, ids.h.new)

std.h.old <- cleaning_MQ(mq.df = mq.h, remove.contaminants = FALSE, remove.decoys = TRUE,
                   poi = NULL, spikeIn = c("P00761","P07477"))
std.h.new <- cleaning_MQ(mq.df = mq.h.new, remove.contaminants = FALSE, remove.decoys = TRUE,
                   poi = NULL, spikeIn = c("P00761","P07477"))
std.y <- cleaning_MQ(mq.df = mq1, remove.contaminants = FALSE, remove.decoys = TRUE,
                   poi = NULL, spikeIn = c("P00761","P07477"))

std.new <- cleaning_MQ(mq.h.new, remove.contaminants = FALSE, remove.decoys = TRUE, poi = NULL, spikeIn = "P07477")
stdinf.new <- extract_proteinID(std.new$Protein.IDs, regex = c(".*(P07477).*", "\\1"), label = "id", routine = NULL)
std.raw.intensity <- MQ_to_longFormat(std.new, y = "raw.intensity",return.dt = TRUE,stdinf.new)


```

### Testing data.base stuff

```{r checkDB, eval = FALSE}
database.name <- "~/Desktop/proteomics2.db"
filename <- system.file("extdata","test_data", "proteinGroups_100mM_new.txt", package = "DepLab")
expt.id <- "pg_300mM"

## create the database, and completely overwrite if it already exists (useful for debugging!)
initialize.database(database.name, organism.name = "yeast", force = TRUE)

## read the data in and turn it into a format suitable for the data base
x <- read.MQ.data(filename, expt.id, organism = "yeast")
head(x)

## testing the manual input of protein IDs ( aka standards aka spike-ins)
y <- read.MQ.data(filename, expt.id, organism = NULL, data.subset = c("P07477", "Q0140","YAL003W"))
head(y)

## come up with some metadata
origin_df <- data.frame(expt_id = expt.id, experimenter = "myself", genotype = "unknown", cell_type = "yeast_cells", harvest_date = "Nov 2016", buffer_composition = "TrisHCl", lysis_method = "standard", digestion_enzyme = "Trypsin", notes = NA)

## optional meta data

msmethods_df <- data.frame(expt_id = expt.id, instrument_id = "X0000",
                           run_date = "Oct 2016", method_length = 1)
dataproc_df <- data.frame(expt_id = expt.id, processing_platform = "unknown", 
                          search_algorithm = "unknown", filtering_algorithm = "unknown",
                          filtering_stringency = "unknown")
prefractionation_df <- data.frame(expt_id = expt.id, column_id = "x",
                                  amount_protein_loaded = 1, sample_vol_loaded = 1, 
                                  lc_flow_rate = 1, lc_fraction_size = 1, 
                                  time_per_fraction = 1, fractions_collected = 1)

## add the data to the db
add.expt.to.database(database.name, data.frame(expt_id = expt.id, organism = "yeast"), 
                     prot.data = NULL, frac.data = x, std.data = y, origin.data = origin_df, 
                     prefractionation.data = NULL, msmethod.data = NULL, dataproc.data = NULL) # NULL could be replaced by the df's generated above
```

### Extracting data from data.base

```{r}
testdat <- as.data.frame(DepLab:::query.measurements.v2("~/Documents/Projects/2015-11_NoahDephoure_Proteomics/fromNoah_2016-03/proteomics.db", "MCF10A_dounce_tsk" , c("HS90A", "HS90B"), "raw.intensity"))
```

### Normalization

```{r checkNorm}
mq.y.1 <- read.MQ.data(filename = system.file("extdata", "test_data", "proteinGroups_100mM_new.txt", package = "DepLab"), expt.id = "100mM", data.subset = "poi", organism = "yeast")
mq.y.3 <- read.MQ.data(filename = system.file("extdata", "test_data", "proteinGroups_300mM_new.txt", package = "DepLab"), expt.id = "300mM", data.subset = "poi", organism = "yeast")
mqcombi <- rbind(mq.y.1, mq.y.3)

smu <- superSmooth_values(long.df = subset(mqcombi, measurement == "raw.intensity"), prot.identifier = "gene_symbol")
fraction.norm <- normalize_values(long.df = smu,
                                  norm.type = "fraction", prot.identifier = "gene_symbol")


y.std.1 <- read.MQ.data(filename = system.file("extdata", "test_data", "proteinGroups_100mM_new.txt", package = "DepLab"), expt.id = "100mM", data.subset = "trypsin", organism = NULL)
y.std.3 <- read.MQ.data(filename = system.file("extdata", "test_data", "proteinGroups_300mM_new.txt", package = "DepLab"), expt.id = "300mM", data.subset = "trypsin", organism = NULL)
std.combi <- rbind(y.std.1, y.std.3)

std.norm <- normalize_values(long.df = subset(mqcombi, measurement == "raw.intensity"), norm.type = "spike-in", prot.identifier = "gene_symbol", std.df = subset(std.combi, measurement == "raw.intensity"))

```

### Plotting
```{r plotting}
mqcombi.plot <- subset(mqcombi, measurement == "raw.intensity" & gene_symbol %in% c("YAL003W (EFB1)", "YAL005C (SSA1)"))
P <- plot_profile(mqcombi.plot, what = c("gene_symbol","expt_id"), color.by = "gene_symbol", split.by = "gene_symbol", line.smooth = TRUE)
print(P)

# plot normalized values
mqcombi.plot.norm.frac <- subset(fraction.norm, gene_symbol %in% c("YAL016W (TPD3)", "YAL005C (SSA1)", "YAL003W (EFB1)"))
plot_profile(mqcombi.plot.norm.frac, what = c("gene_symbol","expt_id"), color.by = "gene_symbol", split.by = "gene_symbol", line.smooth = FALSE)

mqcombi.plot.norm.std <- subset(std.norm, gene_symbol %in% c("YAL016W (TPD3)", "YAL005C (SSA1)", "YAL003W (EFB1)") )
plot_profile(mqcombi.plot.norm.std, what = c("gene_symbol","expt_id"), color.by = "expt_id", split.by = "gene_symbol", line.smooth = FALSE)


# adding marker for molecular weight
mwmark <- data.frame(expt_id = c("100mM","300mM"), MWmarker = c(15,25))
P + geom_vline(data = mwmark, aes(xintercept = MWmarker), linetype="dashed")

# smoothing line instead of geom_line
ggplot(subset(fraction.norm, grepl("YAL005C", gene_symbol)), aes(x=fraction, y=value, colour=expt_id)) + 
  geom_smooth(span = 0.3, se = TRUE) +
  geom_point()+ theme_bw() +
  facet_grid(.~expt_id)
```

### Testing installation

To allow for the installation of the package via `install_url`, we provide a bundled version of the code.
This is an intermediate state between the complete source package and a binary bundle that one would typically download from CRAN.

To **build** the non-binary bundle, first set your working directory to the package folder, then use `devtools::build()` with default settings.

Make sure that `inst/extdata/path_to_custom_complexes.txt` and `inst/extdata/path_to_db.txt` are empty!

```{r buildBundle, eval=FALSE}
# example path
##setwd("/Users/frd2007/Documents/Projects/2015-11_NoahDephoure_Proteomics/dephourelab/DepLab")
devtools::load_all()
# update date and version number in DESCRIPTION, then run document() to update the vignettes and NAMESPACE:
devtools::document()
devtools::build(path = "~/Documents/Projects/2015-11_NoahDephoure_Proteomics/")
```

If `build()` finished successfully, it will tell you where it stored the tarball.
This should be uploaded to `chagall.med.cornell.edu/deplab-pcp`.

To install the package, one can have the working directory pointing anywhere. 
I recommend to **restart your R session** though.

```{r installBundledPackage, eval = FALSE}
setwd("~/Desktop")

# download package bundle and install it
devtools::install_url(url = "http://chagall.med.cornell.edu/deplab-pcp/DepLab_0.1.1.9000.tar.gz")

# attach the package to your current workspace
library(DepLab)

# test if app runs
runPCP()
```

### Adding new libraries

If you change the code in `DepLab.R` or `sql.R` or `shiny_functions.R` and you
need a library that's not yet listed in the DESCRIPTION file of the package, 
do **not** add the usual `library(NewLibrary)` command to the `.R` file.
Instead, follow one of these strategies:

1. **If you need only few functions from the package, e.g. `melt()` from `reshape2`:**
  * add the pkg name to the _Imports:_ section within the DESCRIPTION file
    - this will make sure that the package is installed along with `DepLab`, but
      the functions will not be accessible without loading them
  * in your function, call the external function explicitly via `pkg::fun()`, e.g.
    `reshape2::melt()`
  
2. **If you need many functions from the package, but only within that single function:**
  * add the package name to the _Imports:_ section within the DESCRIPTION file
  * add `roxygen2` roclets for importing the functions that you need
    immediately before your function: `#' @importFrom pkg fun`
      - if you don't want to spell out each function individually, you can also 
        write `#' @import pkg`
  * in your function, you can now call the functions of that package without `pkg::`
  
3. **If you need many functions from the package in multiple new functions of yours:**
  * add the package name to the _Depends:_ section within the DESCRIPTION file
      - this will attach the package and make all functions accessible for the entire `DepLab` package
